### 2019-06-20 test_analytic_geodesic_distance.Rmd
scms now uses silverman's rule but keeps isotropic kernel bandwidth matrix using sigma_min of cosmic reconstruction paper.
added scenario 4

---
title: Test if analytic derivation of geodesic distance for curves from sim_functional_data.R
  is correct
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
library(reticulate)

use_python('/usr/local/bin/python3.7',required=TRUE)
# use_python('/Users/UQAM/anaconda3/bin/python',required=TRUE)
# use_python('/Users/suswei/anaconda3/bin/python',required=TRUE)

```

Check with py_config()

```{r}
py_config()
```

```{r}
source('sim_functional_data.R')
```

The function sim_functional_data.R allows specification of several simulatin scenarios. The samplesize should be set very high so that Floyd's algorithm works well on the noiseless data. The number of time grid points K also should be high so that the vector data really represents the functional curve.

```{r}
# Generate data
sim <- sim_functional_data(sce=4,samplesize=1500,K=150,plot_true=1)
```

The output actually contains more than just the simulated functional data. In particular, it contains the in-sample pairwise geodesic distance matrix (analytic_geo) via analytic derivation. The purpose of this notebook is to see if our analytic derivations match numerical results.

```{r}
names(sim)
```

We would like to compare sim$analytic_geo to geodesic distance estimation via Floyd's algorithm for the noiseless (discretised) curve. This algorithm is guaranteed to do well if we sample the manifold densely enough? (That's why we set the samplesize to be quite high initially.)

```{r}
get_min_num_neighbors = import_from_path("get_min_num_neighbors",path='.')
getIsomapGdist = import_from_path("getIsomapGdist",path='.')

attach(sim)
a = reg_grid[1]
b = tail(reg_grid,1)
K = length(reg_grid)
samplesize = dim(noiseless_data)[1]
noiseless_data_tmp = (sqrt((b-a)/K))*noiseless_data

# Find a grid of possible values for the number of neigbors
num_neigh_min=get_min_num_neighbors$get_min_num_neighbors(noiseless_data_tmp)
num_neigh_true=seq(num_neigh_min,samplesize/2,by=2)

# Calculate the geo matrix for each number of neighbors and keep the one that gives the minimal error
norm_analytic_geo = sqrt(sum(analytic_geo^2))
Error_true_mani_K= rep(0,length(num_neigh_true))
for(j in 1:length(num_neigh_true)){
  IsomapGdist = getIsomapGdist$getIsomapGdist(noiseless_data_tmp,num_neigh_true[j])
  Error_true_mani_K[j]=sqrt(sum((IsomapGdist - analytic_geo )^2))/norm_analytic_geo
}
ind_op_true=min(which(Error_true_mani_K==min(Error_true_mani_K)))
estim_geo_noiseless_data = getIsomapGdist$getIsomapGdist(noiseless_data_tmp,num_neigh_true[ind_op_true])
image.plot(estim_geo_noiseless_data,main='geo estimation')
```

If the analytic derviation is correct, then estim_geo_noiseless_data should be extremely similar to analytic_geo. We can assess their closeness using assess_goodness_estimation.R. Three assessment metrics are returned. If the calculation is correct, rmse should be close to zero and epsilon_isometry_auc and pearson_corr should both be close to 1. Note that running this notebook for sce=1 gives rmse on the order of e-16.
```{r}
library(MESS)
source('assess_goodness_estimation.R')
assess_goodness_estimation(estim_geo_noiseless_data,sim$analytic_geo)
```