### 2019-08-09 index.Rmd
more writing

---
title: "The hidden manifold distance for functional data"
output:
  pdf_document:
    keep_tex: yes
    number_sections: yes
  word_document: default
  html_document:
    df_print: paged
    number_sections: yes
bibliography: references.bib
---

\newcommand {\To}{\rightarrow}
\newcommand {\TO}{\Rightarrow}
\newcommand {\R}{\mathbb{R}}
\newcommand {\Prob}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand {\cov}{\textrm{Cov}}
\newcommand {\var}{\textrm{Var}}
\newcommand {\1}{\textrm{\textbf{1}}}
\newcommand{\M}{\mathcal{M}}

# Abstract
The clustering and classification of discretely-sampled data varying over a continuu, e.g.\ time or space, is an important machine learning task. Data of this nature are also known as functional data.
While one can treat each data point as a vector in Euclidean space, the successful development of functional data analysis demonstrates that it worthwhile to take the functional perspective.
Specifically, functional data analysis is a branch of statistics dedicated to the analysis of infinite-dimensional \textit{smooth} curves. The challenge of analyzing infinite-dimensional objects is ameliorated by the fact that the curves are smooth since the dimensionality is then only artifically high. 
This insight underpins functional data analysis.
There is also another, much less explored, sense in which the dimensionality of the curves under study may be low -- the curves may actually live on a submanifold with low intrinsic dimension.
Interesting classes of functional manifold data include classes of probability density functions and classes of warped curves of a common template function. 
<!-- Since these spaces are not vector spaces, even basic operations such as addition and subtraction require special consideration.-->
This work addresses the estimation of the pairwise geodesic distance between functional manifold data, observed with noise due to both measurement error and the fact that observations may live near, rather than on, the manifold.
<!-- This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.  -->
<!-- We will show that estimation of the geodesic distance cannot easily be accomplished by a naive method such as smoothing followed by a shortest-path algorithm. -->
The proposed methodology first projects the functional data onto the hidden manifold, estimated using subspace-constrained mean-shift. Geodesic distances are subsequently calculated by employing shortest-path algorithms on this estimated manifold. 
Improved estimation of the pairwise geodesic distance has beneficial implications for many downstream tasks in functional data analysis which we illustrate in the case of distanced-based functional classification.

# Introduction 
Many statistical and machine learning methods rely on some measure of distance. 
When the observations are functions, it is especially important to use an appropriate measure. 
Functional data analysis, the study of smooth infinite-dimensional curves, is no exception. Certain classes of functional data naturally live on manifolds, e.g.\ classes of probability density functions and classes of warped curves of a common template function.
In situations where the manifold hypothesis may be plausible, one might consider using a geodesic distance rather than the $L^2$ distance. The benefits of adopting the geodesic distance may be realized in downstream tasks such as distanced-based clustering and classification of functional obvervations.

Specifically, let $X_1,\ldots,X_n$ be a sample of $n$ independant realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$. Suppose additionally that the function $X$ belongs to a Riemannian manifold $\M \subset L^2([a,b],\R)$ with Riemannian metric tensor $g$ which can be used to assign a metric on the manifold as follows [@Lin2014]. For each point $X$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_X$ on the tangent space $T_X \M$. The norm of a tangent vector $V \in T_X \M$ is defined as $$||V|| = \sqrt{g_X(V,V)}.$$ The geodesic distance between two functions $X_i,X_j$ on the manifold $\M$, based on this metric tensor $g$, is defined as
$$ d_{\M}(X_i,X_j):=\inf \{l(\gamma): \gamma \text{ is piecewise-smooth function from } [a,b] \text{ to } \M, \gamma(a) = X_i, \gamma(b) = X_j \}$$
where $$ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt$$
is the length of the curve $\gamma$.

<!-- ???insert example that shows pairwise geodesic distances give visually more interesting results than pairwise $L_2$ distance.??? -->

<!-- (gives examples and references for functional manifolds). -->
<!-- Before continuing, some diambiguities are called for. Contrast existing work on functional data where either the domain is a manifold or the range is a manifold, or both.  -->

<!-- There exist many nonlinear dimension reduction methods for manifolds embedded in Euclidean space, also known as manifold learning methods, which are particularly popular in computer vision. However, the success of these methods typically require the data to be observed with a high signal-to-noise ratio. (is it true? ref?)  -->

Given perfect functional observations $X_i$ and $X_j$, i.e.\ observed on a very dense domain grid with no measurement error, the estimation of the geodesic distance $d_g(X_i,X_j)$ is straightfoward, e.g.\ a shortest-path algorithm such as Floyd-Warhsall [@Floyd1962] may be used.
However, this approach criticaly assumes observations live on the manifold and is thus likely to fail for functional data which are almost certainly observed with noise.
<!-- A preprocessing step where the nosiy functional manifold data is smoothed is also likely to fail. This is because smoothing techniques in functional data analysis are designed for the $L_2$ recovery of the original curve, not to ensure the recovered functional versions of the data lie on or close to the functional manifold $\M$. This presents challenges to the application of many manifold learning techniques which assume fully observed nearly-noiseless inputs. -->
<!-- Estimating the manifold $\M$ from a sample of functions observed discretely and with measurements is a challenging task (results depend on curvature of the manifold, need a lot of data). Even after recovering a functional version of the data, the chances are pretty high that $\tilde X_1,\ldots,\tilde X_n$ do not lie exactly on $\M$.  -->
In this work, we put forth a technique for the specific task of recovering the $n{\times}n$ matrix $G$ of pairwise geodesic distances, i.e.\
$$
G(i,j)=G(j,i) = \begin{cases} 
      d_{\M} (X_i,X_j) & \textrm{if $i\neq j$,} \\
      0 & \textrm{otherwise.}
   \end{cases}
$$ 
In lieu of $X_i$ and $X_j$, we have access only to discretely-observed noisy functional observations $Y_i$ and $Y_j$ that possibly live off the true manifold.

<!-- Pairwise distances are important for many downstream tasks, e.g.  classification, clustering, manifold learning (ISOMAP), nonparametric regression (kernel regression). Certainly, pairwise geodesic distances are legitimate objects of interest in themselves but our motivation is more general in nature. Specifically, we use pairwise distances as an illuminating example to advocate and promote the nonlinear perspective in FDA.  -->

The work of [@ChenMuller2012] was among the first in functional data analysis to consider the manifold hypothesis for functional data. A notion of the mean and variation of functional manifold data was introduced there. Of particular relevance to this work is their proposed P-ISOMAP procedure which adds a data-adaptive penalty to allow for noisy functional observations whereas the classic ISOMAP algorithm [@Tenenbaum2000] assumes observations lie exactly on the manifold. This very drawback of ISOMAP led [@Dimeglio2014] to propose the robust-ISOMAP, a version of ISOMAP that is less sensitive to outliers. We will discuss P-ISOMAP and robust-ISOMAP in more details in the simulation section.

<!-- [@LinYao2017]  -->

# Proposed method for estimating geodesic distances
Suppose that each curve $X_i$ is observed with measurements errors on a grid $T_i=(t_{i1},\ldots,t_{iK})$, i.e. we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$, where the random variables $\epsilon_{ij}$ have mean zero and are uncorrelated with each other. We assume that each observation grid $T_1,\ldots,T_n$ is dense ???add reference for dense???.

We begin by converting the discrete functional observations into continuous ones. Let $\tilde X_1,\ldots,\tilde X_n$ denote the functional versions of the raw data obtained by some smoothing method. Our proposed methodology for estimating the pairwise geodesic distances $\{ d_\M (X_i,X_j)\}_{i>j}$ is agnostic to the smoothing method employed. For example we may employ spline smoothing ???add reference for spline smoothing??? to recover the functional versions of the raw data, i.e.\
\begin{equation}\label{eq_spline_smoothing} 
\tilde X_i = \arg \min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\},
\end{equation}
where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.

Before introducing our procedure, we need to first review the ISOMAP algorithm [@Tenenbaum2000], a three-step procedure that takes as input a set of points $x_1,\ldots,x_n$ in a submanifold of $\R^D$ and produces an embedding in the space $\R^d$ with $d<D$ that preserves pairwise geodesic distances. The procedure is as follows.
\begin{enumerate}
\item Construct a weighted graph $G$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$. Two nodes $x_i$ and $x_j$ are connected by an edge $e_{ij}$ with weight $d_{ij} 1\{d_{ij} \le \epsilon \}$ where $d_{ij}=||x_i-x_j ||_{2}$ and $\epsilon > 0$.
\item Estimate the pairwise geodesic distances $d_g(x_i,x_j)$ based on $G$ using shortest-path algorithms. Specifically, the geodesic distance between two nodes is estimated to be the length of the shortest path in $G$ between these two nodes, i.e.\ the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm [@Floyd1962] or with the Dijkstra algorithm [@Dijkstra59anote].
<!-- Which one do we actually use in our implementation, Floyd-Warshall or Dijkstra's? -->
\item Use multidimensional scaling ???add reference??? to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.
\end{enumerate}
Note that shortest-path algorithms such as the Floyd-Warshall algorithm are used to find the smallest path given a weigthed graph but they do not produce the weighted graphs themselves. Formulating the graph in step 1 of ISOMAP is in fact the most challenging asepct of the algorithm since it depends heavily on the choice of the tuning parameter $\epsilon$.  Both P-ISOMAP and robust ISOMAP are modifications of ISOMAP in the sense that they change the construction of the weighted graph in Step 1.  
<!-- (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, P-ISOMAP is one of these). -->

In what follows, we call IsoGeo the procedure which performs a modified version of the first step of ISOMAP, proposed in @Dimeglio2014, followed by the original second step of ISOMAP. The first step of IsoGeo modifies Step 1 of ISOMAP so that the constructed graph $G$ is independent of the tuning parameter $\epsilon$:
1'. Construct the complete weigthed graph $G_c$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$. Obtain the minimal spanning tree $G_s$ associated with $G_c$ and denote its set of edges by $E_s$. The graph $G$ is obtained by adding all edges $e_{ij}$ to $E_s$ for which the following condition is true:
$$ \overline{x_ix_j} \subset \bigcup_{i=1}^n B(x_i,\epsilon_i),$$
where $B(x_i,\epsilon_i)$ is the open ball of center $x_i$ and radius $\epsilon_i = \max_{e_{ij}\in E_s}d_{ij}$, and $\overline{x_ix_j}= \{x\in R^D \ | \ \exists \lambda \in [0,1], x=\lambda x_i + (1-\lambda)x_j \}$.
Note that IsoGeo has no tuning parameters.

Our method is based on the idea that the underlying functional manifold can be sufficiently well-recovered by the subspace-constrained mean-shift algorithm [@Ozertem2011]. Theoretical justificaton for this can be found in [@Genovese2014]. We expect the points $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n$ to lie close to the real manifold $\M$ and then $d_{\hat \M}$ to be close to $d_\M$. Our procedure is described by the following steps:

<!-- 1. [Susan: I think the smoothing step should be independent of our proposed methodology] Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing: -->
<!-- $$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$ -->
<!-- where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.  -->
\begin{enumerate}
\item Discretise the functions $\tilde X_1,\ldots,\tilde X_n$ by evaluating them on a common grid $\tilde T=(t_{1},\ldots,t_{K})$. Let $s$ be a positive integer, smaller than $K$. Obtain $\tilde X^s_1,\ldots,\tilde X^s_n \in \mathbb R^s$ using multidimensional scaling whereby the pairwise $\mathbb R^K$ Euclidean distances on the discretised functions $\tilde X_i$ are preserved.
\item Apply the subspace constrained mean-shift algorithm [@Ozertem2011] to each of $\tilde X^s_i$ to obtain  $\tilde X^{s,\hat \M}_i$.
\item Use IsoGeo to approximate the pairwise geodesic distances $\{d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$.
\end{enumerate}
Thus our geodesic distance estimator is the $n{\times}n$ matrix $\hat G$ whose elements are given by
$$
\hat G(i,j)=\hat G(j,i) = \left\{ \begin{array}{ll}
 d_{\hat \M}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j) & \textrm{if $i\neq j$,}\\
 0 & \textrm{otherwise.}
  \end{array} \right.
$$
Since ridge estimation suffers from the curse of dimensionality, we first reduce the dimension of our data with multidimensional scaling in Step 1 before applying the mean-shift algorithm.

## Selection of tuning parameters
Describe heuristics. We pick the bandwidth $h$ in the subspace constrained mean-shift using Equation (A1) of [@ChenHo2015].

<!-- ##############  -->

# Simulation study

We perform a simulation study to ascertain the efficacy of our method for estimating pairwise geodesic distances when one only has access to discretely-observed noisy functional data. Three different metrics are used to assess the quality of a pairwise geodesic distance estimator.

* the near isometry metric: this is the area under the receiver operating curve with $\epsilon$ on the $x$-axis and the degree to which near-$\epsilon$ isometry holds on the $y$-axis. Near-$\epsilon$ isometry is a relaxation of isometry measured by the percentage of estimated pairwise distances between $1-\epsilon$ and $1+\epsilon$ of the truth pairwise distance.  
* the relative Frobenius metric: this is given by
$||d-\hat d||/||d||$
where $d$ and $\hat d$ are the true and estimated $n{\times}n$ geodesic distance matrices, respectively. 
* the Pearson correlation metric: this is the Pearson correlation coefficient between the upper diagonal of $d$ and $\hat d$.

## Alternative estimators of geodesic distance [Marie]
We compare the performance of our method to the one of the three following methods:

1.  **Raw Data (RD)** The naive approach consisting in appling directly IsoGeo on the raw vectors $Y_1,\ldots,Y_n \in R^K$. <!-- to obtain an estimator $\hat G_{\textrm{RD}}$.--> Note that this procedure only makes sense if all the grids $T_1,\ldots,T_n$ are the same.
2.  **Spline Smoothing (SS)** The natural method consisting in applying IsoGeo on the smoothed version $\tilde X_1,\ldots,\tilde X_n$ of the raw data obtained by spline smoothing as described in (\ref{eq_spline_smoothing}). <!--Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing. Apply IsoGeo on the vectors $\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n$, where $t_1,\ldots,t_K$ is a regular grid of $[a,b]$, to obtain an estimator $\hat G_{\textrm{SS}}$.-->
3.  **P-ISOMAP (pI)** The two step procedure developed in @ChenMuller2012 where the first step is a modified version of step I of ISOMAP incorporating a penalty in order to robustify the construction of the weighted graph and the second step is step II of ISOMAP.
<!--* **(Random Projection) RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.-->

We ran Chen and Muller's P-ISOMAP separately in Matlab and tuned for the number of neighbors and the penalty parameter using the relative Frobenius assessment metric. We found that the penalty parameter chosen in this way was always zero, thus reducing hte P-ISOMAP procedure to standard ISOMAP. Given this, we did not make further comparison to P-ISOMAP in our simulation study.

## Simulation scenarios 
We shall consider three functional manifolds in our simulation study. As of yet, there is little work as to how to sample from a functional manifold. Even in the Euclidean case, it is not obvious how sampling should be done [@Diaconis2013]. How to properly sample from a functional manifold could be interesting future work. For now, to safeguard against sampling unevenly on the functional manifold, we sample on a very concentrated measure for the intrinsic parameters.

### Isometric functional manifold of normal density functions
<!-- Scenario 2 in sim_functional_data -->
This scenario is modified from what is referred to as Manifold 2 in [@ChenMuller2012] by fixing the variance of the normal density to be $1$. We have $$\M =  \left \{X_\beta: \beta \in [-1,1], t \in [a,b]\right \}$$ with the $L_2$ inner product as the metric tensor of $\M$, where $X_\beta: [a,b] \to \R$ is given by $X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}$. We sample $\beta$ according to a truncated standard normal with support on $[-1,1]$.  We set $a=-4$ and $b=4$.The geodesic distance between the curves $X_{\beta_1}$ and $X_{\beta_2}$ is given by
\begin{eqnarray*}
d(X_{\beta_1},X_{\beta_2}) &=& \int_{\beta_1}^{\beta_2} \left \| \frac{d X_\beta (t)}{d\beta} \right\|_{L^2} d\beta \\
&=&  \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4 \frac{1}{\sqrt{\pi}} \exp\{-(t-\beta)^2\}(t-\beta)^2 dt  }  \  d\beta \\
&=&   \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4(t-\beta)^2 f(t) dt  }  \  d\beta, \textrm{ where $f$ is the density of a N$(\beta,1/2)$ }   \\
&\approx&  \int_{\beta_1}^{\beta_2}  \sqrt{ \frac{1}{2\sqrt{\pi}} \frac{1}{2}} \ d\beta \\
&=& (\beta_2-\beta_1) \frac{1}{2\pi^{1/4}},
\end{eqnarray*}
where the approximation comes from the fact that we are integrating on $[a,b]=[-4,4]$ and not on $\R$. 
We can see this manifold is isometric, since the geodesic distance between $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ is the Euclidan distance between the $\beta$'s, up to some scaling factor. Note that the "straight" line connecting $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ does not always stay inside of $\M$, so we cannot employ the calculation technique of Scenario 1.

### Functional manifold of square root velocity functions
It was shown in [@Joshi2007] that the square root representation of probability density functions has a nice closed form geodesic.
They consider the manifold $$ \M = \{ \psi:[0,1] \to \R : \psi \ge 0, \int_0^1 \psi^2(s) \,ds = 1 \}$$ with the metric tensor given by the Fisher-Rao metric tensor $$ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds $$ for two tangent vectors $v_1,v_2 \in T_\psi(\M)$. Note that this concides with the $L_2[0,1]$ inner product. [@Joshi2007] showed that the geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply $$d(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.$$
We will specifically examine the square root of $Beta(\alpha,\beta)$ distributions which is supported on $[0,1]$. That is, $$ M = \{ \psi_{\alpha,\beta}: 1 \le \alpha \le 5, 2 \le \beta \le 5\} $$ where $\psi_{\alpha,\beta}: [0,1] \to \R$ is the pdf of $Beta(\alpha,\beta)$. We sample $\alpha$ according to a truncated normal with mean 3 and variance 0.09 with support on $[1,5]$. We sample $\beta$ according to a truncated normal with mean 3.5 and variance 0.09 with support on $[2,5]$.

### Functional manifold of warping functions
<!-- Scenario 5 -->
This is based on Equations (17) and (18) of [@Kneip2008] but with $z_{i1}, z_{i2}$ set to $1$. (Equation 17 has a typo where the exponentials are missing negative signs).Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\} + \exp\{(t+1.5)^2/2\}$$ and $$ h_\alpha(t) = 6 \frac{ \exp\{\alpha(t+3)/6\} - 1}{\exp\{\alpha\}-1}, \alpha \ne 0 $$ and $h_\alpha(t) = t$ if $\alpha = 0$. 
Consider the manifold $$ M = \{X_\alpha: -1 \le \alpha \le 1\} $$. 
We sample $\alpha$ according to a truncated standard normal with support on $[-1,1]$.
The geodesic distance is then $$ d(X_{\alpha_1},X_{\alpha_2}) = \int_{\alpha_1}^{\alpha_2} \left \| \frac{d X_\alpha (t)}{d\alpha} \right\|_{L^2} d\alpha.$$

## Results
Submit compare_methods.sh as slurm job using sbatch.
```{r, engine = 'bash', eval = FALSE}
sbatch compare_methods.sh
```

Simulation results are then visualized by invoking cluster_results.R. There's some unfortunate hardcoding still in cluster_results.R
```{r, eval=FALSE}
R cluster_results.R
```
<!--

## The cost of employing proposed method when manifold is flat [Susan/Marie?]

Suppose we take scenario 1 where the geodesic distance coincides with the $L_2$ distance. Then what is the cost of employing our distance estimator compared to smoothing the data and doing numerical integration to find the $L_2$ distance?

-->

# Distance-based functional classification [Marie]

In this section, we explore whether our geodesic distance estimator has benefits for downstream analysis task. There are many tasks we could consider here such as distance-based nonparametric regression and distanced-based functional clustering, but we will focus on distance-based functional classification. It must be noted that while curve alignment, also known as curve registration, is necessarily performed as a preprocessing technique prior to clustering and classification, our geodesic distance estimator allows one to forsake this step.

For simplicity, assume the task is binary classification. Associated to each functional object $x$ is a binary $y$ indicating class membership.
Consider the classifier proposed in Ferraty and Vieu (2003,2006) which is a functional version of the Nadaraya-Watson kernel estimator of class membership probabilities:
$$
\hat p(y = 0 | x) \frac{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] 1(y_i = 0) }{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] }
$$
We shall compare our method to using $L_2$ distance, possibly weighted, and with curve registration already accomplished. Describe alternative methods in detail.

The bandwidth in the classifier should be tuned individually for each method. Also we might need to tune MDS dimension $s$ since in real data, the dimension of the manifold might be much higher than encountered in the simulation scenarios where it never goes above 2.

Datasets used by functional classification papers
\begin{itemize}
\item Wheat, rainfall and phoneme in Aurore's paper "Achieving near-perfect classification for functional data"
\item Berkeley growth curves in [@ChenReiss2014].
\item Tecator and phoneme in [@Galeano2015] Mahalanobis technometrics paper.
\item yeast cell cycle gene expression (can't find this publicly) in [@LengMuller2005] "Classification using functional data analysis for temporal gene expression data"
\end{itemize}

Datasets used in functional manifold papers
\begin{itemize}
\item Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in [@ChenMuller2012]
\item Tecator in [@LinYao2017] contamination paper
\item Berkeley growth, gait cycle in [@Dimeglio2014] robust isomap paper
\end{itemize}

<!-- ## Distance-based functional clustering -->

<!-- Describe the $k$-medoids clustering method. -->

<!-- ### Simulated data -->

<!-- ###  Real data -->

# References