### 2019-08-24 index.Rmd
cleaned up intro a bit

---
title: "The hidden manifold distance for functional data"
output: 
  html_document:
    fig_caption: true
bibliography: references.bib
---

```{r, include=FALSE}
source("cluster_results.R",echo=FALSE)
```

\newcommand {\To}{\rightarrow}
\newcommand {\TO}{\Rightarrow}
\newcommand {\R}{\mathbb{R}}
\newcommand {\Prob}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand {\cov}{\textrm{Cov}}
\newcommand {\var}{\textrm{Var}}
\newcommand {\1}{\textrm{\textbf{1}}}
\newcommand{\M}{\mathcal{M}}

# Abstract
The clustering and classification of functional data, discretely-sampled curves varying over a continuum, is frequently encountered in machine learning.
While one can naively treat the data as vectors in Euclidean space, one may do better to regard the data as functions.
For certain types of functional data such as probability density functions or warped curves of a common template function, there is further structure to exploit. Namely, the curves may lie on a manifold of low intrinsic dimension.
This work addresses the estimation of pairwise geodesic distances between noisy functional manifold data that live near, rather than on, a manifold.
This setting cannot be tackled by classic manifold learning techniques which require data to live exactly on a manifold.
<!-- We will show that estimation of the geodesic distance cannot easily be accomplished by a naive method such as smoothing followed by a shortest-path algorithm. -->
The proposed methodology first sends the observed functional data to the hidden manifold, estimated using subspace-constrained mean-shift.
Geodesic distances are subsequently calculated by employing shortest-path algorithms on this estimated manifold. 
Improved estimation of the pairwise geodesic distance has beneficial implications for many downstream tasks in functional data analysis which we illustrate in the case of distance-based functional classification/clustering.

# Introduction 
Many statistical and machine learning methods rely on some measure of distance. For example, clustering and classification of functional data, discretely-sampled curves varying over a continuum, is a common machine learning task in which distance plays a crucial role. In a naive treatment, the data may be simply processed as vectors in Euclidean space. Doing so however risks losing the richness of functional data; indeed, one may do better to regard the data as functions.

The challenge of analyzing infinite-dimensional objects can be ameliorated if the curves are smooth since the dimensionality is then only artificially high.
This insight underpins functional data analysis, a subfield of statistics that studies functional data. This has many implications for the processing of functional data. 
For example, the $L^2$ distance between the smoothed functional observations has advantages over the Euclidean distance between functional observations treated as vectors.

In this work, we advocate that for functional manifold data, i.e. functional data that lie on a manifold, the geodesic distance is potentially even better than the $L^2$ distance, with benefits realized in downstream tasks such as distanced-based clustering and classification.
In other words, when the manifold hypothesis is plausible, e.g.\ for classes of probability density functions and classes of warped curves of a common template function, clustering and classification using the goedesic distance may give better results than using the $L_2$ distance. The manifold setting presents its own challenges since for nonlinear manifolds, even basic operations such as addition and subtraction require special consideration.

Let $X_1,\ldots,X_n$ be a sample of $n$ independant realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$. Suppose additionally that the function $X$ belongs to a Riemannian manifold $\M \subset L^2([a,b],\R)$ with Riemannian metric tensor $g$ which can be used to assign a metric on the manifold as follows [@Lin2014]. For each point $X$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_X$ on the tangent space $T_X \M$. The norm of a tangent vector $V \in T_X \M$ is defined as $$||V|| = \sqrt{g_X(V,V)}.$$ The geodesic distance between two functions $X_i,X_j$ on the manifold $\M$, based on this metric tensor $g$, is defined as
$$ d_{\M}(X_i,X_j):=\inf_{\gamma:[a,b] \to \M, \gamma(a) = X_i, \gamma(b) = X_j} l(\gamma)$$
where $$ l(\gamma) := \int_a^b || \frac{\,d\gamma}{\,d t}(t) || \,dt$$
is the length of the piecewise-smooth curve $\gamma$. 
<!-- ???Is it clear how $l(\gamma)$ depends on $g$??? -->

<!-- ???insert example that shows pairwise geodesic distances give visually more interesting results than pairwise $L_2$ distance.??? -->

<!-- (gives examples and references for functional manifolds). -->
<!-- Before continuing, some diambiguities are called for. Contrast existing work on functional data where either the domain is a manifold or the range is a manifold, or both.  -->

<!-- There exist many nonlinear dimension reduction methods for manifolds embedded in Euclidean space, also known as manifold learning methods, which are particularly popular in computer vision. However, the success of these methods typically require the data to be observed with a high signal-to-noise ratio. (is it true? ref?)  -->

Ideally, the functions $X_i$ and $X_j$ are observed on a very dense domain grid with no measurement error. Then the estimation of the geodesic distance $d_\M(X_i,X_j)$ is straightfoward, e.g.\ a shortest-path algorithm such as Floyd-Warhsall [@Floyd1962] may be used.
However, most shortest-path algorithms critically assume observations live on the manifold [@Yeh2005] and are thus likely to fail for functional data which are almost certainly observed with noise.
<!-- A preprocessing step where the nosiy functional manifold data is smoothed is also likely to fail. This is because smoothing techniques in functional data analysis are designed for the $L_2$ recovery of the original curve, not to ensure the recovered functional versions of the data lie on or close to the functional manifold $\M$. This presents challenges to the application of many manifold learning techniques which assume fully observed nearly-noiseless inputs. -->
<!-- Estimating the manifold $\M$ from a sample of functions observed discretely and with measurements is a challenging task (results depend on curvature of the manifold, need a lot of data). Even after recovering a functional version of the data, the chances are pretty high that $\tilde X_1,\ldots,\tilde X_n$ do not lie exactly on $\M$.  -->
In this work, we put forth a technique for the recovery of the $n{\times}n$ matrix $G$ of pairwise geodesic distances, i.e.\
$$
G(i,j)=G(j,i) = \begin{cases} 
      d_{\M} (X_i,X_j) & \textrm{if $i\neq j$,} \\
      0 & \textrm{otherwise.}
   \end{cases}
$$ 
In lieu of $X_i$ and $X_j$, we have access only to discretely-sampled noisy functional observations $Y_i$ and $Y_j$ that possibly lie off the true manifold.

<!-- Pairwise distances are important for many downstream tasks, e.g.  classification, clustering, manifold learning (ISOMAP), nonparametric regression (kernel regression). Certainly, pairwise geodesic distances are legitimate objects of interest in themselves but our motivation is more general in nature. Specifically, we use pairwise distances as an illuminating example to advocate and promote the nonlinear perspective in FDA.  -->

The work of [@ChenMuller2012] was among the first in functional data analysis to consider the manifold hypothesis for functional data. A notion of the mean and variation of functional manifold data was introduced there. Of particular relevance to this work is their proposed P-ISOMAP procedure which adds a penalty to allow for noisy functional observations whereas the classic ISOMAP algorithm [@Tenenbaum2000] assumes observations lie exactly on the manifold.  [@Dimeglio2014] also realized this drawback to ISOMAP and proposed a procedure we will call robust-ISOMAP that is less sensitive to outliers. We will discuss P-ISOMAP and robust-ISOMAP in more details in the simulation section.

<!-- [@LinYao2017]  -->

# Proposed method for estimating geodesic distances 
Suppose that each curve $X_i \in (\M,g)$ is observed with measurement error on a time grid $T_i=(t_{i1},\ldots,t_{iK}), a \le t_{i1} < \ldots < t_{iK} \le b$, i.e.\ we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$, where the random variables $\epsilon_{ij}$ have mean zero and are uncorrelated with each other. We assume that each observation grid $T_1,\ldots,T_n$ is dense, i.e.\ $K$ is large.

We begin by converting the discretly-sampled functional observations into continuous ones. Let $\tilde X_1,\ldots,\tilde X_n$ denote the functional versions of the raw data obtained by some smoothing method. For example, we may employ spline smoothing [@Ramsay2005] to recover the functional versions of the raw data, i.e.\
\begin{equation}\label{eq_spline_smoothing} 
\tilde X_i = \arg \min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\},
\end{equation} where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$. The proposed methodology for estimating the pairwise geodesic distances $\{ d_\M (X_i,X_j)\}_{i>j}$ is agnostic to the smoothing method employed.

Our method is based on the idea that the underlying functional manifold $\M$ can be sufficiently well-recovered by the subspace-constrained mean-shift (SCMS) algorithm [@Ozertem2011]. Let $\hat M \subset L^2([a,b],\mathbb R)$ denote the collection of basins of attraction of the estimated probability density function of $X$. The SCMS algorithm sends each point $\tilde X_i \in L^2([a,b],\mathbb R)$ to a point in $\hat M$. This destination is unique and we denote it $\tilde X_i^{\hat \M}$. We expect the points $\tilde X^{\hat \M}_1,\ldots,\tilde X^{\hat \M}_n \in \hat M$ to lie close to the real manifold $\M$. Theoretical justificaton that $\hat M$ as estimated by SCMS is a reasonable surrogate for the true manifold $\M$ can be found in [@Genovese2014].

Before presenting our procedure, we need to first review the ISOMAP algorithm [@Tenenbaum2000], a three-step procedure that takes as input a set of points $x_1,\ldots,x_n$ in a submanifold $\M$ of $\R^D$ and produces an embedding in the space $\R^d$ with $d<D$ that preserves pairwise geodesic distances. The procedure is as follows.

1. Construct a weighted graph $G$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$. Two nodes $x_i$ and $x_j$ are connected by an edge $e_{ij}$ with weight $d_{ij} 1\{d_{ij} \le \epsilon \}$ where $d_{ij}=||x_i-x_j ||_{2}$ and $\epsilon > 0$.
2. Estimate the pairwise geodesic distances $d_\M(x_i,x_j)$ based on $G$ using shortest-path algorithms. Specifically, the geodesic distance between two nodes is estimated to be the length of the shortest path in $G$ between these two nodes, i.e.\ the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm [@Floyd1962] or with the Dijkstra algorithm [@Dijkstra59anote].
3. Use multidimensional scaling to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.

Note that shortest-path algorithms such as the Floyd-Warshall algorithm are used to find the smallest path \textit{given} a weigthed graph but they do not produce the weighted graphs themselves. Both P-ISOMAP and robust-ISOMAP are modifications of ISOMAP in the sense that they change the construction of the weighted graph in step 1.  
<!-- (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, P-ISOMAP is one of these). -->

In what follows, we call IsoGeo the two-step procedure which performs a modified version of the first step of ISOMAP followed by the original second step of ISOMAP. The first step of IsoGeo constructs the weighted graph $G$ accoding to @Dimeglio2014 so that graph is independent of the tuning parameter $\epsilon$. Specifically, we begin by constructing the complete ???explain what complete means??? weighted graph $G_c$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$. Next, we obtain the minimal spanning tree associated with $G_c$ and denote its set of edges by $E_s$. The graph $G$ is obtained by adding all edges $e_{ij}$ to $E_s$ for which the following condition is true:
$$ \overline{x_ix_j} \subset \bigcup_{i=1}^n B(x_i,\epsilon_i),$$
where $B(x_i,\epsilon_i)$ is the open ball of center $x_i$ and radius $\epsilon_i = \max_{e_{ij}\in E_s}d_{ij}$, and $\overline{x_ix_j}= \{x\in R^D \ | \ \exists \lambda \in [0,1], x=\lambda x_i + (1-\lambda)x_j \}$. Let the distance estimated by IsoGeo be denoted $\text{IsoGeo}(x_i,x_j)$.
Note that IsoGeo has no tuning parameters.

We are now ready to describe our procedure. Since SCMS is based on kernel density estimation, we first reduce the dimension of our data with multidimensional scaling before applying SCMS to avoid the curse of dimensionality.

<!-- 1. [Susan: I think the smoothing step should be independent of our proposed methodology] Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing: -->
<!-- $$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$ -->
<!-- where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.  -->
1. Discretise the functions $\tilde X_1,\ldots,\tilde X_n$ by evaluating them on a dense common grid $\tilde T=(t_{1},\ldots,t_{D})$. Let $s$ be a positive integer, much smaller than $D$. Obtain $\tilde X^s_1,\ldots,\tilde X^s_n \in \mathbb R^s$ using multidimensional scaling whereby the pairwise $\mathbb R^D$ Euclidean distances of the discretised functions $\tilde X_i$ are preserved.
2. Apply the subspace constrained mean-shift algorithm [@Ozertem2011] to each of $\tilde X^s_i$ to obtain $\tilde X^{s,\hat \M}_i$.
3. Use IsoGeo on the inputs $(\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n)$ to obtain $\text{IsoGeo}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$.

Our geodesic distance estimator is the $n{\times}n$ matrix $\hat G$ whose elements are given by
$$
\hat G(i,j)=\hat G(j,i) = \left\{ \begin{array}{ll}
 \text{IsoGeo}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j) & \textrm{if $i\neq j$,}\\
 0 & \textrm{otherwise.}
  \end{array} \right.
$$

There are two tuning parameters to our procedure. First, there is the dimension $s$ in step 1. In our simulations we try $s \in \{1,2,3\}$. The other parameter is the bandwidth $h$ in the subspace constrained mean-shift. In our simulations we set $h$ according to the heuristic in Equation (A1) of [@ChenHo2015]. In reality, depending on the downstream task, both $s$ and $h$ can be tuned in a data-adaptive way, say using cross-validation.

<!-- ##############  -->

# Simulation study

We perform a simulation study to ascertain the efficacy of our method for estimating pairwise geodesic distances. Three different metrics are used to assess the quality of a pairwise geodesic distance estimator.

The first metric assesses near-$\epsilon$ isometry, which we define as the percentage of estimated pairwise geodesic distances between $1-\epsilon$ and $1+\epsilon$ of the corresponding true pairwise geodesic distance. We form the receiver operating curve with $\epsilon$ on the $x$-axis and near-$\epsilon$ isometry on the $y$-axis. The metric is the area under the receiver operating curve.

The second metric $||d-\hat d||/||d||$ is the Frobenius norm of the estimation error matrix $d-\hat d$, relative to the Frobenius norm of the true distance matrix $d$. The third metric is the Pearson correlation coefficient between the upper diagonal of $d$ and $\hat d$.

## Alternative estimators of geodesic distance [Marie]
We compare the performance of our method to each of the following four methods. The first method, denoted RD for raw data, takes the naive approach of applying IsoGeo directly on the raw vectors $Y_1,\ldots,Y_n \in R^K$. Note that this procedure is only sensible if all the grids $T_1,\ldots,T_n$ are the same. In the second method, we first smooth the discretely-sampled noisy functional data and then estimate their pairwise distance using the $L^2$ distance.

The third alternative, denoted SS for spline smoothing, applies IsoGeo on the smoothed version $\tilde X_1,\ldots,\tilde X_n$ of the raw data obtained by spline smoothing as described in (\ref{eq_spline_smoothing}). <!--Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing. Apply IsoGeo on the vectors $\{(\tilde X_i(t_1), \ldots, X_i(t_K)\}_{i=1}^n$, where $t_1,\ldots,t_K$ is a regular grid of $[a,b]$, to obtain an estimator $\hat G_{\textrm{SS}}$.-->
The final alternative we consider is P-ISOMAP, denoted pI, which is the two-step procedure developed in @ChenMuller2012 designed to handle noisy functional data. In the first step, a penalty is incorporated to robustify the construction of the weighted graph. This is followed by the original second step of ISOMAP.
<!--* **(Random Projection) RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.-->

We ran Chen and Muller's P-ISOMAP separately in Matlab and tuned for the number of neighbors and the penalty parameter using the relative Frobenius assessment metric. We found that the penalty parameter chosen in this way was always zero, thus reducing hte P-ISOMAP procedure to standard ISOMAP. Given this, we did not make further comparison to P-ISOMAP in our simulation study.

## Simulation scenarios 
We shall consider three functional manifolds in our simulation study. As of yet, there is little work as to how to sample from a functional manifold. Even in the Euclidean case, it is not obvious how sampling should be done [@Diaconis2013]. How to properly sample from a functional manifold could be interesting future work. For now, to safeguard against sampling unevenly on the functional manifold, we sample on a very concentrated measure for the intrinsic parameters.

### Isometric functional manifold of normal density functions
<!-- Scenario 2 in sim_functional_data -->
This scenario is modified from what is referred to as Manifold 2 in [@ChenMuller2012] by fixing the variance of the normal density to be $1$. We have $$\M =  \left \{X_\beta: \beta \in [-1,1], t \in [a,b]\right \}$$ with the $L_2$ inner product as the metric tensor of $\M$, where $X_\beta: [a,b] \to \R$ is given by $X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}$. First, note that the "straight" line connecting $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ does not always stay inside of $\M$. Thus the geodesic distance will not coincide with the $L^2$ distance.

We set $a=-4$ and $b=4$ and sample $\beta$ according to a truncated standard normal with support on $[-1,1]$.   The geodesic distance between the curves $X_{\beta_1}$ and $X_{\beta_2}$ is given by
\begin{eqnarray*}
d(X_{\beta_1},X_{\beta_2}) &=& \int_{\beta_1}^{\beta_2} \left \| \frac{d X_\beta (t)}{d\beta} \right\|_{L^2} d\beta \\
% &=&  \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4 \frac{1}{\sqrt{\pi}} \exp\{-(t-\beta)^2\}(t-\beta)^2 dt  }  \  d\beta \\ 
&=&   \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4(t-\beta)^2 f(t) dt  }  \  d\beta   \\
&\approx&  \int_{\beta_1}^{\beta_2}  \sqrt{ \frac{1}{2\sqrt{\pi}} \frac{1}{2}} \ d\beta \\
&=& (\beta_2-\beta_1) \frac{1}{2\pi^{1/4}}
\end{eqnarray*} 
where $f$ is the density of a N$(\beta,1/2)$ and the approximation in the next to last step comes from the fact that we are integrating on $[a,b]=[-4,4]$ and not on $\R$. 
We can see this manifold is isometric, since the geodesic distance between $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ is proportional to $\beta_2-\beta_1$, the Euclidean distance between the $\beta$'s.

### Functional manifold of square root velocity functions
<!-- Scenario 4 in sim_functional_data -->
It was shown in @Joshi2007 that the square root representation of probability density functions gives rise to a simple closed-form geodesic distance.
They consider the manifold $$ \M = \{ \psi:[0,1] \to \R : \psi \ge 0, \int_0^1 \psi^2(s) \,ds = 1 \}$$ with the metric tensor given by the Fisher-Rao metric tensor $$ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds $$ for two tangent vectors $v_1,v_2 \in T_\psi(\M)$. Note that this concides with the $L_2[0,1]$ inner product. @Joshi2007 showed that the geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply $$d_\M(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.$$
We will specifically examine the square root of $Beta(\alpha,\beta)$ distributions which is supported on $[0,1]$. That is, $$ M = \{ \psi_{\alpha,\beta}: 1 \le \alpha \le 5, 2 \le \beta \le 5\} $$ where $\psi_{\alpha,\beta}: [0,1] \to \R$ is the pdf of $Beta(\alpha,\beta)$. We sample $\alpha$ according to a truncated normal with mean 3 and variance 0.09 with support on $[1,5]$. We sample $\beta$ according to a truncated normal with mean 3.5 and variance 0.09 with support on $[2,5]$.

### Functional manifold of warping functions
<!-- Scenario 5 in sim_functional_data -->
This is based on Equations 17 and 18 in @Kneip2008 but with $z_{i1}, z_{i2}$ both set to $1$. (Note that Equation 17 has a typo where the exponentials are missing negative signs.) Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\} + \exp\{(t+1.5)^2/2\}$$ and $$ h_\alpha(t) = 6 \frac{ \exp\{\alpha(t+3)/6\} - 1}{\exp\{\alpha\}-1}$$ if $\alpha \ne 0$ and $h_\alpha(t) = t$ otherwise.

Consider the manifold $$ M = \{X_\alpha: -1 \le \alpha \le 1\}.$$ 
We sample $\alpha$ according to a truncated standard normal with support on $[-1,1]$.
The geodesic distance is then $$ d(X_{\alpha_1},X_{\alpha_2}) = \int_{\alpha_1}^{\alpha_2} \left \| \frac{d X_\alpha (t)}{d\alpha} \right\|_{L^2} d\alpha$$ which is estimated using numerical integration.

## Results

The tuning parameters for the proposed method were chosen as described at the end of Section \ref{proposed-method-for-estimating-geodesic-distances}. In step 1 of our procedure, we take $D = 100$ for the size of the common time grid. We will investigate how two parameters affect the performance of the various methods in consideration. The first one is the signal-to-noise ratio ???describe??? which we fix at either $0.1$ (low) or $0.5$ (high). The second parameter is $K$, the dimension of the dscrietly-observed noisy $Y_i$'s. We set $K$ to either $30$ (sparse) or $100$ (abundant).

We show the performance of the SS and L2 method versus our method for three different values of $s$. We omit the result of the square root velocity scenario because all methods perform comparably. We can see from Figures \ref{fig:normal} and \ref{fig:warping} that ???.

```{r, engine = 'bash', eval = FALSE, include=FALSE}
sbatch compare_methods.sh
scp susanwei@spartan.hpc.unimelb.edu.au:/home/susanwei/functional_manifold_data/taskid* /Documents/Githuib/functional_manifold_data/
```

```{r, echo=FALSE, eval=FaALSE, results='hide'}

setwd("./sim_results")

method_names=c("SS","s=1","s=2","s=3","L2")
nr_methods = length(method_names)

combination_res_assess<- cbind(rep(1:nr_methods,3),rep(1:3,each=nr_methods)) # nr_methods*3 combinations of methods and assesment measures
# TODO: fix hardcoding
combination_para <- expand.grid(c(5,2,4),c(0.1,0.5),c(30,100)) # combinations of the parameters, copied from compare_methods.R

apply(combination_para,1,create_boxplot,mat_ind=combination_res_assess,method_names=method_names,nr_methods=nr_methods)

setwd("../")
```

\begin{figure}
\centering
\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_MSE}
\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_isometry}
\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_Pearson}
\caption{Normal density}
\label{fig:normal}
\end{figure}

<!-- \begin{figure} -->
<!-- \includegraphics{sim_results/sce=4_SNR=high_Kobs=30_MSE} -->
<!-- \includegraphics{sim_results/sce=4_SNR=high_Kobs=30_isometry} -->
<!-- \includegraphics{sim_results/sce=4_SNR=high_Kobs=30_Pearson} -->
<!-- \caption{Square root velocity} -->
<!-- \end{figure} -->

\begin{figure}
\centering
\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_MSE}
\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_isometry}
\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_Pearson}
\caption{Warping function}
\label{fig:warping}
\end{figure}

<!--

## The cost of employing proposed method when manifold is flat [Susan/Marie?]

Suppose we take scenario 1 where the geodesic distance coincides with the $L_2$ distance. Then what is the cost of employing our distance estimator compared to smoothing the data and doing numerical integration to find the $L_2$ distance?

-->

# Distance-based functional classification [Marie]

In this section, we explore whether our geodesic distance estimator has benefits for downstream analysis task. There are many tasks we could consider here such as distance-based nonparametric regression and distanced-based functional clustering, but we will focus on distance-based functional classification. It must be noted that while curve alignment, also known as curve registration, is necessarily performed as a preprocessing technique prior to clustering and classification, our geodesic distance estimator allows one to forsake this step.

For simplicity, assume the task is binary classification. Associated to each functional object $x$ is a binary $y$ indicating class membership.
Consider the classifier proposed in Ferraty and Vieu (2003,2006) which is a functional version of the Nadaraya-Watson kernel estimator of class membership probabilities:
$$
\hat p(y = 0 | x) \frac{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] 1(y_i = 0) }{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] }
$$
We shall compare our method to using $L_2$ distance, possibly weighted, and with curve registration already accomplished. Describe alternative methods in detail.

The bandwidth in the classifier should be tuned individually for each method. Also we might need to tune MDS dimension $s$ since in real data, the dimension of the manifold might be much higher than encountered in the simulation scenarios where it never goes above 2.

Datasets used by functional classification papers
\begin{itemize}
\item Wheat, rainfall and phoneme in Aurore's paper "Achieving near-perfect classification for functional data"
\item Berkeley growth curves in [@ChenReiss2014].
\item Tecator and phoneme in [@Galeano2015] Mahalanobis technometrics paper.
\item yeast cell cycle gene expression (can't find this publicly) in [@LengMuller2005] "Classification using functional data analysis for temporal gene expression data"
\end{itemize}

Datasets used in functional manifold papers
\begin{itemize}
\item Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in [@ChenMuller2012]
\item Tecator in [@LinYao2017] contamination paper
\item Berkeley growth, gait cycle in [@Dimeglio2014] robust isomap paper
\end{itemize}

<!-- ## Distance-based functional clustering -->

<!-- Describe the $k$-medoids clustering method. -->

<!-- ### Simulated data -->

<!-- ###  Real data -->

# References