### 2019-08-30 index.Rmd

---
title: "The hidden manifold distance for functional data"
output:
  html_document:
    fig_caption: yes
---

```{r, include=FALSE}
source("cluster_results.R",echo=FALSE)
```

\newcommand {\To}{\rightarrow}
\newcommand {\TO}{\Rightarrow}
\newcommand {\R}{\mathbb{R}}
\newcommand {\Prob}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand {\cov}{\textrm{Cov}}
\newcommand {\var}{\textrm{Var}}
\newcommand {\1}{\textrm{\textbf{1}}}
\newcommand{\M}{\mathcal{M}}

# Introduction 
Many statistical and machine learning methods rely on some measure of distance. For example, clustering and classification of functional data, discretely-sampled curves varying over a continuum, is a common learning task in which distance plays a crucial role. In a naive treatment, the data may be simply processed as vectors in Euclidean space. Doing so however risks losing the richness of functional data; indeed, one may do better to regard the data as functions. The challenge of analyzing infinite-dimensional objects can be ameliorated if the curves are smooth since the dimensionality is then only artificially high.
This insight underpins functional data analysis, a subfield of statistics that studies functional data.

<!-- This has many implications for the processing of functional data.  -->
<!-- For example, the $L^2$ distance between the smoothed functional observations has advantages over the Euclidean distance between noisy functional observations treated as vectors.  -->

In this work, we advocate that for functional manifold data, i.e. functional data that lie on a manifold, the geodesic distance is more appropriate than the $L^2$ distance.
Indeed, when the manifold hypothesis is plausible, e.g.\ for classes of probability density functions and classes of warped curves of a common template function, clustering and classification using the geodesic distance may give better results than using the $L^2$ distance.

The manifold setting presents certainly challenges since for nonlinear manifolds, even basic operations such as addition and subtraction require special consideration. We will carefully lay out a technique for estimating the geodesic distance between noisy functional manifold data.

Let $X_1,\ldots,X_n$ be a sample of $n$ independent realizations of a random variable $X$ that takes value in the Hilbert space $L^2([a,b],\R)$. Suppose additionally that the function $X$ belongs to a Riemannian manifold $\M \subset L^2([a,b],\R)$ with Riemannian metric tensor $g$ which can be used to assign a metric on the manifold as follows \cite{Lin2014}. For each point $X$ on the manifold, the Riemannian metric tensor $g$ has an inner product $g_X$ on the tangent space $T_X \M$. The norm of a tangent vector $V \in T_X \M$ is defined as $$||V|| = \sqrt{g_X(V,V)}.$$ The geodesic distance between two functions $X_i,X_j$ on the manifold $\M$, based on this metric tensor $g$, is defined as
$$ d_{\M}(X_i,X_j):=\inf_{\gamma:[0,1] \to \M, \gamma(0) = X_i, \gamma(1) = X_j} l(\gamma)$$
where $$ l(\gamma) := \int_0^1 || \frac{\,d\gamma}{\,d t}(t) || \,dt$$
is the length of the piecewise-smooth curve $\gamma$. 
<!-- ???Is it clear how $l(\gamma)$ depends on $g$??? -->

<!-- ???insert example that shows pairwise geodesic distances give visually more interesting results than pairwise $L^2$ distance.??? -->

<!-- (gives examples and references for functional manifolds). -->
<!-- Before continuing, some diambiguities are called for. Contrast existing work on functional data where either the domain is a manifold or the range is a manifold, or both.  -->

<!-- There exist many nonlinear dimension reduction methods for manifolds embedded in Euclidean space, also known as manifold learning methods, which are particularly popular in computer vision. However, the success of these methods typically require the data to be observed with a high signal-to-noise ratio. (is it true? ref?)  -->

Ideally, the functions $X_i$ and $X_j$ are observed on a very dense domain grid with no measurement error. Then the geodesic distance $d_\M(X_i,X_j)$ can be estimated using any of a variety of shortest-path algorithms, e.g.\ the Floyd-Warhsall \cite{Floyd1962} algorithm.
However, most shortest-path algorithms critically assume observations live near the manifold with little noise \cite{Yeh2005} and are thus likely to fail for functional data which can be observed with low signal-to-noise ratio.
<!-- A preprocessing step where the nosiy functional manifold data is smoothed is also likely to fail. This is because smoothing techniques in functional data analysis are designed for the $L^2$ recovery of the original curve, not to ensure the recovered functional versions of the data lie on or close to the functional manifold $\M$. This presents challenges to the application of many manifold learning techniques which assume fully observed nearly-noiseless inputs. -->
<!-- Estimating the manifold $\M$ from a sample of functions observed discretely and with measurements is a challenging task (results depend on curvature of the manifold, need a lot of data). Even after recovering a functional version of the data, the chances are pretty high that $\tilde X_1,\ldots,\tilde X_n$ do not lie exactly on $\M$.  -->
In this work, we put forth a technique for the recovery of the $n{\times}n$ matrix $G$ of pairwise geodesic distances, i.e.\
$$
G(i,j)=G(j,i) = \begin{cases} 
      d_{\M} (X_i,X_j) & \textrm{if $i\neq j$,} \\
      0 & \textrm{otherwise.}
   \end{cases}
$$ 
In lieu of $X_i$ and $X_j$, we have access only to discretely-sampled noisy functional observations $Y_i$ and $Y_j$ that possibly lie off the true manifold.

<!-- Pairwise distances are important for many downstream tasks, e.g.  classification, clustering, manifold learning (ISOMAP), nonparametric regression (kernel regression). Certainly, pairwise geodesic distances are legitimate objects of interest in themselves but our motivation is more general in nature. Specifically, we use pairwise distances as an illuminating example to advocate and promote the nonlinear perspective in FDA.  -->

The work of \cite{ChenMuller2012} was among the first in functional data analysis to consider the manifold hypothesis for functional data. A notion of the mean and variation of functional manifold data was introduced there. Of particular relevance to this work is their proposed P-ISOMAP procedure which allows for noisy functional observations. This is in contrast to the classic ISOMAP algorithm \cite{Tenenbaum2000} which assumes observations lie exactly on the manifold.  \cite{Dimeglio2014} also realized this drawback to ISOMAP and proposed a procedure we will call robust-ISOMAP for functional data that is less sensitive to outliers. We will discuss P-ISOMAP and robust-ISOMAP in more details below.

<!-- \cite{LinYao2017]  -->

# Proposed method for estimating geodesic distances 
Suppose that each curve $X_i \in (\M,g)$ is observed with measurement error on a time grid $T_i=(t_{i1},\ldots,t_{iK}), a \le t_{i1} < \ldots < t_{iK} \le b$, i.e.\ we observe a sample of $K$-dimensional vectors $Y_1,\ldots,Y_n$ with $Y_{ij} = X_i(t_{ij}) + \epsilon_{ij}$. Let the random variables $\epsilon_{ij}$ have mean zero and be uncorrelated with each other. We assume that each observation grid $T_1,\ldots,T_n$ is dense, i.e.\ $K$ is large.

We begin by converting the discretely-sampled functional observations into continuous ones. Let $\tilde X_1,\ldots,\tilde X_n$ denote the functional versions of the raw data obtained by some smoothing method. For example, we may employ spline smoothing \cite{Ramsay2005} to recover the functional versions of the raw data, i.e.\
\begin{equation}\label{eq_spline_smoothing} 
\tilde X_i = \arg \min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\},
\end{equation} 
where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$. 
<!-- The proposed methodology for estimating the pairwise geodesic distances $\{ d_\M (X_i,X_j)\}_{i>j}$ is agnostic to the smoothing method employed. -->
Our method is based on the idea that the underlying functional manifold $\M$ can be sufficiently well-recovered by the subspace-constrained mean-shift (SCMS) algorithm \cite{Ozertem2011} which we now review briefly in the Euclidean case. Suppose $Z_1,\ldots,Z_n$ is a sample of realisations of a random vector $Z \in \M$. We can use a kernel density estimator $\hat f$ based on this data to estimate the probability density function of $Z$. 
Let the set $\hat M$ be comprised of the basins of attraction of $\hat f$.
The SCMS algorithm sends each point $Z_i$ to its destination in $\hat M$.
Theoretical justification that $\hat M$ as estimated by SCMS is a reasonable surrogate for the true manifold $\M$ can be found in \cite{Genovese2014}.

Before presenting our procedure, we also need to review the ISOMAP algorithm \cite{Tenenbaum2000}, a three-step procedure that takes as input a set of points $x_1,\ldots,x_n$ in a submanifold $\M \subset \R^D$ and produces an embedding in the space $\R^d$ with $d<D$ that preserves pairwise geodesic distances. The procedure is as follows.

1. Construct a weighted graph $G$ with nodes corresponding to the observations $x_1,\ldots,x_n\in \R^D$. Two nodes $x_i$ and $x_j$ are connected by an edge $e_{ij}$ with weight $d_{ij} 1\{d_{ij} \le \epsilon \}$ where $d_{ij}=||x_i-x_j ||_{2}$ and $\epsilon > 0$.
2. Estimate the pairwise geodesic distances $d_\M(x_i,x_j)$ based on $G$ using shortest-path algorithms. Specifically, this is the sum of the weights of the edges forming the shortest path, which is calculated either with the Floyd-Warshall algorithm \cite{Floyd1962} or with the Dijkstra algorithm \cite{Dijkstra59anote}.
3. Use multidimensional scaling to obtain an embedding in $\R^d$ that preserves the pairwise geodesic distances estimated above.

Note that shortest-path algorithms such as the Floyd-Warshall algorithm are used to find the smallest path \textit{given} a weighted graph but they do not produce the weighted graphs themselves. Both P-ISOMAP and robust-ISOMAP are modifications of ISOMAP in the sense that they change the construction of the weighted graph in step 1 above.
<!-- (which is the difficult part, depend on number of neighbors, many method out there to calculate that graph in a robust way, P-ISOMAP is one of these). -->

In what follows, we call IsoGeo the two-step procedure which performs a modified version of the first step of ISOMAP followed by the original second step of ISOMAP. The first step of IsoGeo constructs the complete weighted ???Marie: does this mean all weights are nonzero??? graph $G_c$ according to \cite{Dimeglio2014} so that the graph is independent of the tuning parameter $\epsilon$. Next, we obtain the minimal spanning tree associated with $G_c$ and denote its set of edges by $E_s$. The graph $G$ is obtained by adding all edges $e_{ij}$ to $E_s$ for which the following condition is true:
$$ \overline{x_ix_j} \subset \bigcup_{i=1}^n B(x_i,\epsilon_i),$$
where $B(x_i,\epsilon_i)$ is the open ball of center $x_i$ and radius $\epsilon_i = \max_{e_{ij}\in E_s}d_{ij}$, and $\overline{x_ix_j}= \{x\in R^D \ | \ \exists \lambda \in [0,1], x=\lambda x_i + (1-\lambda)x_j \}$. Let the distance estimated by IsoGeo be denoted $\text{IsoGeo}(x_i,x_j)$.
Note that IsoGeo has no tuning parameters.

We are now ready to describe our procedure. Since SCMS is based on kernel density estimation, we first reduce the dimension of our data with multidimensional scaling before applying SCMS to avoid the curse of dimensionality.

<!-- 1. [Susan: I think the smoothing step should be independent of our proposed methodology] Transform each vector $Y_i$ into a function $\tilde X_i$ by spline smoothing: -->
<!-- $$ \tilde X_i = \arg\min_{f\in C^2[0,1]}\left\{\sum_{j=1}^{K}\left(f(t_{ij})-Y_{ij}\right)^2+\lambda \|\partial^2_tf\|^2_{L^2}\right\}$$ -->
<!-- where $\lambda>0$ is a tuning parameter controlling the smoothness of $\tilde X_i$.  -->
1. Discretise the functions $\tilde X_1,\ldots,\tilde X_n$ by evaluating them on a dense common grid $\tilde T=(t_{1},\ldots,t_{D})$. Let $s$ be a positive integer, much smaller than $D$. Obtain $\tilde X^s_1,\ldots,\tilde X^s_n \in \mathbb R^s$ using multidimensional scaling so that the pairwise $\mathbb R^D$ Euclidean distances of the discretised functions $\tilde X_i$ can be preserved as much as possible.
2. Apply the subspace constrained mean-shift algorithm \cite{Ozertem2011} to each of $\tilde X^s_i$ to obtain its destination $\tilde X^{s,\hat \M}_i$.
3. Use IsoGeo on the inputs $\tilde X^{s,\hat \M}_1,\ldots,\tilde X^{s,\hat \M}_n \in \R^s$ to obtain $\text{IsoGeo}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j)\}_{i>j}$.
Our geodesic distance estimator is the $n{\times}n$ matrix $\hat G$ whose elements are given by
$$
\hat G(i,j)=\hat G(j,i) = \left\{ \begin{array}{ll}
 \text{IsoGeo}(\tilde X^{s,\hat \M}_i,\tilde X^{s,\hat \M}_j) & \textrm{if $i\neq j$,}\\
 0 & \textrm{otherwise.}
  \end{array} \right.
$$

There are two tuning parameters to our procedure. First, there is the dimension $s$ in step 1. In our simulations we try $s \in \{1,2,3\}$. The other parameter is the bandwidth $h$ in the subspace constrained mean-shift. In our simulations we set $h$ according to the heuristic in Equation (A1) of \cite{ChenHo2015}. In reality, depending on the downstream task, both $s$ and $h$ can be tuned in a data-adaptive way, say using cross-validation.

<!-- ##############  -->

# Simulation study

We perform simulations to study various estimators of pairwise geodesic distances. We compare the performance of our method to each of the following four alternatives. The first alternative, denoted RD, takes the naive approach of applying IsoGeo directly on the raw vectors $Y_1,\ldots,Y_n \in R^K$. Note that this procedure is only sensible if all the time grids $T_1,\ldots,T_n$ are the same.

For the next two alternatives, we first smooth the discretely-sampled noisy functional data using splines as described in (\ref{eq_spline_smoothing}). Then, we either estimate pairwise distance between the smoothed $\tilde X_1,\ldots,\tilde X_n$ evaluated on a common time grid of $D$ points using IsoGeo (denoted Spline IsoGeo) or the $L^2$ distance (denoted Spline L2).

The final alternative we consider is P-ISOMAP which is the two-step procedure developed in \cite{ChenMuller2012} for handling noisy functional data. In the first step, a penalty is incorporated to robustify the construction of the weighted graph. This is followed by the original second step of ISOMAP.
<!--* **(Random Projection) RP**  same method as our but change step 2 by : obtain $s$-dimensional representation by random projection and setp 4 by obtain $\hat G$ using a ensemble method.-->

We ran P-ISOMAP separately in Matlab and tuned for the number of neighbors and the penalty parameter using the relative Frobenius assessment metric. We found that the penalty parameter chosen in this oracle fashion was always zero, thus reducing the P-ISOMAP procedure to standard ISOMAP. Given this, we did not make further comparison to P-ISOMAP in our simulation study. We also do not present the results for the RD method which always produced poor estimates of the geodesic distance.

Three different metrics are used to assess the quality of a pairwise geodesic distance estimator. The first metric assesses near-$\epsilon$ isometry, which we define as the percentage of estimated pairwise geodesic distances between $1-\epsilon$ and $1+\epsilon$ of the corresponding true pairwise geodesic distance. We form the receiver operating curve with $\epsilon$ on the $x$-axis and near-$\epsilon$ isometry on the $y$-axis. The metric is the area under this receiver operating curve. Thus a good estimator has high values for the near-isometry metric.

The second metric $||G-\hat G||/||G||$ is the Frobenius norm of the estimation error matrix $G-\hat G$, relative to the Frobenius norm of the true distance matrix $G$. The third metric is the Pearson correlation coefficient between the upper diagonal of $G$ and $\hat G$. A good geodesic distance estimator has low Frobenius norm and high Pearson correlation coefficient.

We shall consider three functional manifolds in our simulation study. Throughout, we make use of a helpful property of one-dimensional manifolds. If $X \in \M$, with the $L^2$ metric tensor, depends on a parameter $\theta \in \mathbb R$, then 
$$
d_\M(X_{\theta_1},X_{\theta_2}) = \int_{\theta_1}^{\theta_2} || \frac{\partial X}{\partial \theta}(t)||_{L^2} \,d\theta.
$$

We also need to make a disclaimer about our sampling technique below. Proper sampling from a manifold remains an open problem. Even in the Euclidean case, it is not obvious how sampling should be done \cite{Diaconis2013}. For now, to safeguard against sampling unevenly on the functional manifold, we sample on a very narrow support of the manifold parameters.

## Isometric functional manifold of normal density functions
<!-- Scenario 2 in sim_functional_data -->
We take Manifold 2 in \cite{ChenMuller2012} and set the variance of the normal density to be $1$. Let $X_\beta: [a,b] \to \R$ be given by $X_\beta(t) = \frac{1}{\sqrt{2\pi}} \exp{[-\frac{1}{2}(t-\beta)^2]}$. Consider $$\M = \left \{X_\beta: \beta \in [-1,1], t \in [a,b]\right \}$$ with the $L^2$ inner product as the metric tensor. First, note that the "straight" line connecting $X_{\beta_1}$ and $X_{\beta_2}$ does not always stay inside of $\M$. Thus the geodesic distance will not coincide with the $L^2$ distance.

We set $a=-4$ and $b=4$ and sample $\beta$ according to a truncated standard normal with support on $[-1,1]$. The geodesic distance between the curves $X_{\beta_1}$ and $X_{\beta_2}$ is given by
\begin{eqnarray*}
d_\M(X_{\beta_1},X_{\beta_2}) &=& \int_{\beta_1}^{\beta_2} \left \| \frac{d X_\beta (t)}{d\beta} \right\|_{L^2} d\beta \\
% &=&  \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4 \frac{1}{\sqrt{\pi}} \exp\{-(t-\beta)^2\}(t-\beta)^2 dt  }  \  d\beta \\ 
&=&   \int_{\beta_1}^{\beta_2} \sqrt{ \frac{1}{2\sqrt{\pi}} \int_{-4}^4(t-\beta)^2 f(t) dt  }  \  d\beta   \\
&\approx&  \int_{\beta_1}^{\beta_2}  \sqrt{ \frac{1}{2\sqrt{\pi}} \frac{1}{2}} \ d\beta \\
&=& (\beta_2-\beta_1) \frac{1}{2\pi^{1/4}}
\end{eqnarray*}
\noindent where $f$ is the density of a N$(\beta,1/2)$ and the approximation in the next to last step comes from integrating on $[a,b]=[-4,4]$ rather than $\R$. 
We can see this manifold is isometric, since the geodesic distance between $X_{\beta_1}$ and $X_{\beta_2}$ in $\M$ is proportional to $\beta_2-\beta_1$.

## Functional manifold of square root velocity functions
<!-- Scenario 4 in sim_functional_data -->
It was shown in \cite{Joshi2007} that the square root representation of probability density functions gives rise to a simple closed-form geodesic distance.
They consider the manifold $$ \M = \{ \psi:[0,1] \to \R : \psi \ge 0, \int_0^1 \psi^2(s) \,ds = 1 \}$$ with the Fisher-Rao metric tensor $$ <v_1,v_2> = \int_0^1 v_1(s) v_2(s) \,ds $$ for two tangent vectors $v_1,v_2 \in T_\psi(\M)$. Note that this coincides with the $L^2[0,1]$ inner product. It was shown in \cite{Joshi2007} that the geodesic distance between any two $\psi_1$ and $\psi_2$ in $\M$ is simply $$d_\M(\psi_1,\psi_2) = \cos^{-1}<\psi_1,\psi_2>.$$
We will specifically examine the square root of $Beta(\alpha,\beta)$ densities, which are of course supported on $[0,1]$. That is, consider $$ M = \{ \sqrt{\psi_{\alpha,\beta}}: 1 \le \alpha \le 5, 2 \le \beta \le 5\} $$ where $\psi_{\alpha,\beta}: [0,1] \to \R$ is the probability density function of a Beta random variable with shape parameters $\alpha$ and $\beta$. We sample $\alpha$ according to a truncated normal with mean 3 and variance 0.09 with support on $[1,5]$. We sample $\beta$ according to a truncated normal with mean 3.5 and variance 0.09 with support on $[2,5]$.

## Functional manifold of warped functions
<!-- Scenario 5 in sim_functional_data -->
This is based on Equations 17 and 18 in \cite{Kneip2008} but with $z_{i1}, z_{i2}$ both set to $1$. (Note that Equation 17 therein has a typo where the exponentials are missing negative signs.) Let $X_\alpha(t) = \mu(h_\alpha(t))$ be defined on $[-3,3]$ where $$ \mu(t) = \exp\{(t-1.5)^2/2\}+\exp\{(t+1.5)^2/2\}$$ and $$ h_\alpha(t) = 6\frac{\exp\{\alpha(t+3)/6\}-1}{\exp\{\alpha\}-1}$$ if $\alpha \ne 0$ and $h_\alpha(t) = t$ otherwise.

Consider the manifold $$ M = \{X_\alpha: -1 \le \alpha \le 1\}.$$ 
We sample $\alpha$ according to a truncated standard normal with support on $[-1,1]$.
The geodesic distance is then $$ d_\M(X_{\alpha_1},X_{\alpha_2}) = \int_{\alpha_1}^{\alpha_2} \left \| \frac{d X_\alpha (t)}{d\alpha} \right\|_{L^2} d\alpha.$$ We estimate this using numerical integration.

## Results

The tuning parameters for the proposed method were chosen as described at the end of the methodology section. In step 1 of our procedure, we take $D = 100$ for the size of the common time grid. We will investigate how two parameters affect the performance of the various methods in consideration. The first one is the signal-to-noise ratio ???[Marie] describe??? which we fix at either $0.1$ (low) or $0.5$ (high). The second parameter is $K$, the dimension of the discretely-observed noisy $Y_i$'s. We set $K$ to either $30$ (sparse) or $100$ (abundant).

We will present the results for the high signal-to-noise and sparse $K$ setting. We show the performance of the Spline IsoGeo and Spline L2 method versus our method for three dimensions $s=1,2,3$. We present the result of the square root velocity manifold scenario in the Appendix. We can see from Figures \ref{fig:normal} and \ref{fig:warping} that spline smoothing followed by either IsoGeo or the $L^2$ distance calculation estimates the pairwise geodesic distances reasonably well though not as well as the proposed method for $s=2$ and $s=3$. Interestingly, when $s$ in our method is not selected properly, for instance when $s=1$, our estimator performs poorly in terms of the relative Frobenius norm although still better than Spline IsoGeo and Spline L2 in the other two metrics. It may be possible that for certain downstream tasks such as clustering or classification, precise estimation of the geodesic distance is less important than one that respects the ordering of the distances. In such cases, the near-isometry and Pearson correlation metrics may give a better indication of downstream performance of the geodesic distance estimator.

```{r, engine = 'bash', eval = FALSE, include=FALSE}
sbatch compare_methods.sh
scp susanwei@spartan.hpc.unimelb.edu.au:/home/susanwei/functional_manifold_data/taskid* /Documents/Githuib/functional_manifold_data/
```

```{r, echo=FALSE, eval=TRUE, results='hide'}

setwd("./sim_results")

method_names=c("Spline IsoGeo","s=1","s=2","s=3","Spline L2")
nr_methods = length(method_names)

combination_res_assess<- cbind(rep(1:nr_methods,3),rep(1:3,each=nr_methods)) # nr_methods*3 combinations of methods and assesment measures
# TODO: fix hardcoding
combination_para <- expand.grid(c(5,2,4),c(0.1,0.5),c(30,100)) # combinations of the parameters, copied from compare_methods.R

apply(combination_para,1,create_boxplot,mat_ind=combination_res_assess,method_names=method_names,nr_methods=nr_methods)

setwd("../")
```

\begin{figure}
\centering
\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_MSE}
\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_isometry}
\includegraphics[height=0.3\textheight]{sim_results/sce=2_SNR=high_Kobs=30_Pearson}
\caption{Manifold of $N(\beta,1)$ probability density functions. Across all three metrics, our method with $s=2$ performs the best, followed by our method with $s=3$.}
\label{fig:normal}
\end{figure}

\begin{figure}
\centering
\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_MSE}
\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_isometry}
\includegraphics[height=0.3\textheight]{sim_results/sce=5_SNR=high_Kobs=30_Pearson}
\caption{Manifold of warped functions. The best performing estimators are given by our method with $s=2$ and then $s=3$.}
\label{fig:warping}
\end{figure}

<!--

## The cost of employing proposed method when manifold is flat [Susan/Marie?]

Suppose we take scenario 1 where the geodesic distance coincides with the $L^2$ distance. Then what is the cost of employing our distance estimator compared to smoothing the data and doing numerical integration to find the $L^2$ distance?

-->

<!-- # Distance-based functional classification [Marie] -->

<!-- In this section, we explore whether our geodesic distance estimator has benefits for downstream analysis task. There are many tasks we could consider here such as distance-based nonparametric regression and distanced-based functional clustering, but we will focus on distance-based functional classification. It must be noted that while curve alignment, also known as curve registration, is necessarily performed as a preprocessing technique prior to clustering and classification, our geodesic distance estimator allows one to forsake this step.  -->

<!-- For simplicity, assume the task is binary classification. Associated to each functional object $x$ is a binary $y$ indicating class membership. -->
<!-- Consider the classifier proposed in \cite{Ferraty2006} which is a functional version of the Nadaraya-Watson kernel estimator of class membership probabilities: -->
<!-- $$ -->
<!-- \hat p(y = 0 | x) \frac{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] 1(y_i = 0) }{ \sum_{i=1}^n K[h^{-1} d(x,x_i)] } -->
<!-- $$ -->
<!-- We shall compare our method to using $L^2$ distance, possibly weighted, and with curve registration already accomplished. Describe alternative methods in detail.  -->

<!-- The bandwidth in the classifier should be tuned individually for each method. Also we might need to tune MDS dimension $s$ since in real data, the dimension of the manifold might be much higher than encountered in the simulation scenarios where it never goes above 2.  -->

<!-- Datasets used by functional classification papers -->
<!-- \begin{itemize} -->
<!-- \item Wheat, rainfall and phoneme in Aurore's paper "Achieving near-perfect classification for functional data" -->
<!-- \item Berkeley growth curves in \cite{ChenReiss2014}. -->
<!-- \item Tecator and phoneme in \cite{Galeano2015} Mahalanobis technometrics paper. -->
<!-- \item yeast cell cycle gene expression (can't find this publicly) in \cite{LengMuller2005} "Classification using functional data analysis for temporal gene expression data" -->
<!-- \end{itemize} -->

<!-- Datasets used in functional manifold papers -->
<!-- \begin{itemize} -->
<!-- \item Berkeley growth, yeast cell cycle gene expression (can't find this publicly) in \cite{ChenMuller2012} -->
<!-- \item Tecator in \cite{LinYao2017} contamination paper -->
<!-- \item Berkeley growth, gait cycle in \cite{Dimeglio2014} robust isomap paper -->
<!-- \end{itemize} -->

<!-- ## Distance-based functional clustering -->

<!-- Describe the $k$-medoids clustering method. -->

<!-- ### Simulated data -->

<!-- ###  Real data -->

<!-- # Acknowledgement -->
<!-- The authors would like to acknowledge support from the ARC Centre of Excellence for Mathematical & Statistical Frontiers (ACEMS) which enabled the two authors to work on the project at the University of Melbourne.  -->

<!-- # Appendices -->
<!-- The result of the square root velocity manifold scenario is presented in Figure \ref{fig:betas}. The Spline L2 method has similar performance to our proposed method. This could be due to the fact that we did not sample the manifold broadly enough so that the sample lives in a locally flat region where the $L^2$ distance holds.  -->

<!-- \begin{figure} -->
<!-- \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_MSE} -->
<!-- \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_isometry} -->
<!-- \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_Pearson} -->
<!-- \label{fig:betas} -->
<!-- \caption{Manifold of square root Beta densities. Besides Spline IsoGeo, all methods perform comparably well.} -->
<!-- \end{figure} -->