### 2019-09-03 paperAAAI.tex

\def\year{2020}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
\usepackage{aaai20}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet} % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{graphicx}  % DO NOT CHANGE THIS
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS

\def\tightlist{}

% 
% 
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
% 
% 
% 
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{amsbsy}
\usepackage{graphics}
\usepackage{epsfig}

%\nocopyright
%PDF Info Is REQUIRED.
% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
 \pdfinfo{
/Title (The hidden manifold distance for functional data)
/Author (Susan Wei and Marie-H\'el\`{e}ne Descary)
} %Leave this	
% /Title ()
% Put your actual complete title (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case
% Leave the space between \Title and the beginning parenthesis alone
% /Author ()
% Put your actual complete list of authors (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case. 
% Each author should be only by a comma. If the name contains accents, remove them. If there are any LaTeX commands, 
% remove them.

% DISALLOWED PACKAGES
% \usepackage{authblk} -- This package is specifically forbidden
% \usepackage{balance} -- This package is specifically forbidden
% \usepackage{caption} -- This package is specifically forbidden
% \usepackage{color (if used in text)
% \usepackage{CJK} -- This package is specifically forbidden
% \usepackage{float} -- This package is specifically forbidden
% \usepackage{flushend} -- This package is specifically forbidden
% \usepackage{fontenc} -- This package is specifically forbidden
% \usepackage{fullpage} -- This package is specifically forbidden
% \usepackage{geometry} -- This package is specifically forbidden
% \usepackage{grffile} -- This package is specifically forbidden
% \usepackage{hyperref} -- This package is specifically forbidden
% \usepackage{navigator} -- This package is specifically forbidden
% (or any other package that embeds links such as navigator or hyperref)
% \indentfirst} -- This package is specifically forbidden
% \layout} -- This package is specifically forbidden
% \multicol} -- This package is specifically forbidden
% \nameref} -- This package is specifically forbidden
% \natbib} -- This package is specifically forbidden -- use the following workaround:
%\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}} 
%\newcommand{\citep}{\cite} 
%\newcommand{\citealp}[1]{\citeauthor{#1} \citeyear{#1}}
% \usepackage{savetrees} -- This package is specifically forbidden
% \usepackage{setspace} -- This package is specifically forbidden
% \usepackage{stfloats} -- This package is specifically forbidden
% \usepackage{tabu} -- This package is specifically forbidden
% \usepackage{titlesec} -- This package is specifically forbidden
% \usepackage{tocbibind} -- This package is specifically forbidden
% \usepackage{ulem} -- This package is specifically forbidden
% \usepackage{wrapfig} -- This package is specifically forbidden
% DISALLOWED COMMANDS
% \nocopyright -- Your paper will not be published if you use this command
% \addtolength -- This command may not be used
% \balance -- This command may not be used
% \baselinestretch -- Your paper will not be published if you use this command
% \clearpage -- No page breaks of any kind may be used for the final version of your paper
% \columnsep -- This command may not be used
% \newpage -- No page breaks of any kind may be used for the final version of your paper
% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
% \pagestyle -- This command may not be used
% \tiny -- This is not an acceptable font size.
% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference

\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.

% The file aaai20.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\setlength\titlebox{2.5in} % If your paper contains an overfull \vbox too high warning at the beginning of the document, use this
% command to correct it. You may not alter the value below 2.5 in
\title{The Hidden Manifold Distance for Functional Data}
%Your title must be in mixed case, not sentence case. 
% That means all verbs (including short verbs like be, is, using,and go), 
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash

% \author{Susan Wei\textsuperscript{\rm 1} and Marie-H\'el\`{e}ne Descary\textsuperscript{\rm 2}\\ % All authors must be
% \textsuperscript{\rm 1} The University of Melbourne, \textsuperscript{\rm 2}Universit\'e du Qu\'ebec \`{a} Montr\'eal\\ }

%in the same font size and format. Use \Large and \textbf to achieve this result when breaking a line

%If you have multiple authors and multiple affiliations
% use superscripts in text and roman font to identify them. For example, Sunil Issar,\textsuperscript{\rm 2} J. Scott Penberthy\textsuperscript{\rm 3} George Ferguson,\textsuperscript{\rm 4} Hans Guesgen\textsuperscript{\rm 5}. Note that the comma should be placed BEFORE the superscript for optimum readability
 % email address must be in roman text type, not monospace or sans serif

\begin{document}

\maketitle

\begin{abstract}
In many machine learning applications, the observed data is a
discretely-sampled realisation of a continuous process. While one can
naively treat such data as vectors in Euclidean space, one may do better
to regard them as functional data.  
When the underlying curves of interest lie on a manifold of low intrinsic
dimension, e.g.~probability density
functions or warped curves of a common template function, there is further structure to exploit. 
This work addresses the estimation of pairwise geodesic
distances between functional manifold data that are observed with noise,
causing them to lie off the manifold. This setting falls outside of
classic manifold learning techniques which require data to live exactly
on or very near a manifold. The proposed methodology first sends the
observed functional data to the hidden manifold, estimated using
subspace-constrained mean-shift. Geodesic distances are subsequently
calculated by employing shortest-path algorithms on this estimated
manifold. Improved estimation of the pairwise geodesic distance has
potential benefits for downstream tasks such as distance-based functional classification.
\end{abstract}

\input{index}

\input{MarieDataAnalysis}

%\section{Acknowledgement}\label{acknowledgement}

%The authors would like to acknowledge support from the ARC Centre of
%Excellence for Mathematical and Statistical Frontiers (ACEMS) which
%enabled the two authors to work on the project at the University of
%Melbourne.

% \section{Appendices}\label{appendices}
% 
% The result of the square root velocity manifold scenario is presented in
% Figure \ref{fig:betas}. The Spline L2 method has similar performance to
% our proposed method. This could be due to the fact that we did not
% sample the manifold broadly enough so that the sample lives in a locally
% flat region where the \(L^2\) distance holds.
% 
% \begin{figure}
% \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_MSE}
% \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_isometry}
% \includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_Pearson}
% \caption{Manifold of square root Beta densities. Besides Spline IsoGeo, all methods perform comparably well.}
% \label{fig:betas}
% \end{figure}

\newpage

\bibliography{references}
\bibliographystyle{aaai}

\end{document}